---
title: "R Notebook"
output: html_notebook
---
```{r echo=FALSE}
library(readr)
library(tidyverse)
library(ggplot2)
library(scales)
library(emmeans)
library(broom) 
library(multcompView)
library(multcomp) 
library(Metrics)
library(caret)
library(FSA)
library(dplyr)
``` 

```{r}
salary_df <- read.csv("Data/datascience.csv")
#salary_data_with_category <- read.csv("Data/datascienceold.csv")  
```

#Left join with salary_data_no_category to impute the job_category
```{r}
salary_df <- salary_df %>%
  filter(company_location == "US") 

split_index <- createDataPartition(salary_df$job_title, p = 0.7, list = FALSE)

train_data <- salary_df[split_index, ]
test_data <- salary_df[-split_index, ]
```


#Pre-process factoring 
```{r}  
salary_data <- subset(train_data, select = - c(company_location,employee_residence,salary_currency,salary))
test_data <- subset(test_data, select = - c(company_location,employee_residence,salary_currency,salary))

cols_to_factor <- c("work_year", "experience_level", "employment_type", "job_title",  "remote_ratio", "company_size")
salary_data[cols_to_factor] <- lapply(salary_data[cols_to_factor], as.factor)
test_data[cols_to_factor] <- lapply(test_data[cols_to_factor], as.factor)
```


#Encoding
```{r} 
dummy <- dummyVars(~ work_year + experience_level + employment_type + remote_ratio + company_size, data = salary_data)
# Apply one-hot encoding to the dataset
salary_data_one_hot <- predict(dummy, newdata = salary_data)
salary_data_one_hot <- data.frame(salary_data_one_hot)
salary_data_combined <- cbind(salary_data, salary_data_one_hot)
salary_data <- salary_data_combined[, !names(salary_data_combined) %in% c("work_year", "experience_level", "employment_type",  "remote_ratio", "company_size")]
```


#Correlation
```{r}
means <- salary_data_cleaned %>%
  group_by(job_title) %>%
  summarise(job_title_encoded = mean(salary_in_usd), .groups = "drop")

salary_data_cor <- salary_data_cleaned %>%
  left_join(means, by = "job_title") %>%
  dplyr::select(-job_title) %>%
  rename(job_title = job_title_encoded)
cor_matrix <- cor(salary_data_cor, use = "complete.obs", method = "pearson")
print(cor_matrix) 
```


#Scaling
```{r}
min_salary <- min(train_data$salary_in_usd)
max_salary <- max(train_data$salary_in_usd)

salary_data <- salary_data %>%
  mutate(salary_in_usd = (salary_in_usd - min_salary) / (max_salary - min_salary))
#view(salary_data)
```

#kruskal.test
```{r} 
salary_data <- salary_data[, sapply(salary_data, function(x) length(unique(x)) > 1)]
 
factor_names <- names(salary_data)[-which(names(salary_data) == "salary_in_usd")]
 
results_list <- lapply(factor_names, function(factor_var) {
  test <- kruskal.test(as.formula(paste("salary_in_usd ~", factor_var)), data = salary_data)
  data.frame(
    Factor = factor_var,
    Statistic = round(test$statistic, 3),
    P_Value = round(test$p.value, 4),
    stringsAsFactors = FALSE
  )
})
 
results_table <- do.call(rbind, results_list) 

print(results_table %>%  arrange(desc(Statistic)))
```
Most of our factors have very low p-values (p < 0.001), which means they have statistically significant effects on salary. Especially job_title and experience Levels they have huge test statistics, meaning there's a strong difference in salary across their groups.

#Comparision between job roles and the salary
```{r}
salary_data$salary_in_usd <- as.numeric(salary_data$salary_in_usd)
salary_data$job_title <- as.factor(salary_data$job_title)

top_jobs <- names(sort(table(salary_data$job_title), decreasing = TRUE)[1:10])
filtered_data <- salary_data[salary_data$job_title %in% top_jobs, ]
dunn_result <- dunnTest(salary_in_usd ~ job_title, data = filtered_data, method = "bonferroni")
print(dunn_result) 
```

there are statistically significant differences in salary between at least some job titles.

# SMOTE
```{r}
max_per_title <- max(table(salary_data$job_title))
salary_data <- salary_data %>%
  group_by(job_title) %>%
  slice(rep(1:n(), length.out = max_per_title)) %>%
  ungroup()
```

#Do the model
```{r}
#source("RandomForest/Model.R")
#source("DecisionTree/Model.R")
source("./LinearRegression/Model.R")
```

#Apply encoding on test and scaling
```{r}
# Apply one-hot encoding to the dataset
test_salary_data_one_hot <- predict(dummy, newdata = test_data)
test_salary_data_one_hot <- data.frame(test_salary_data_one_hot)
test_salary_data_combined <- cbind(test_data, test_salary_data_one_hot)
test_data <- test_salary_data_combined[, !names(test_salary_data_combined) %in% c("work_year", "experience_level", "employment_type",  "remote_ratio", "company_size")]

test_data <- test_data %>%
  mutate(salary_in_usd = (salary_in_usd - min_salary) / (max_salary - min_salary))

```


#Test the model
```{r}
test_data$predicted_salary <- predict(model, newdata = test_data)

test_data$salary_in_usd <- (test_data$salary_in_usd * (max_salary - min_salary)) + min_salary
test_data$predicted_salary_original <- (test_data$predicted_salary * (max_salary - min_salary)) + min_salary
head(test_data[, c("job_title", "salary_in_usd", "predicted_salary_original")])
```

```{r}
# Calculate RMSE and MAE
rmse_value <- rmse(test_data$salary_in_usd, test_data$predicted_salary_original)
mae_value <- mae(test_data$salary_in_usd, test_data$predicted_salary_original)

# Print RMSE and MAE
print(paste("RMSE:", rmse_value))
print(paste("MAE:", mae_value))
```


